Thesis: draft

### 2021-04-30 draft.tex
latest update

  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution. \textbf{(a)} Using the PR filtering algorithm; \textbf{(b)} Using the REC filtering algorithm. The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the average of $10^2$ independent simulations.}

@@

Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion). Figure \ref{PR_frac_compl_stubb_20-80} shows the results for the PR filtering algorithm and Figure \ref{REC_frac_compl_stubb_20-80} shows those for the REC filtering algorithm. The data points are shown together with their standard error $SE$, note that these are very small and almost not visible in the figure. The interpretation of the $y$-axis is slightly different than before. In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both. In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity. We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities. Hence, we opted for a slightly different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) and then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before). Note that by taking the average, we neglect the skewed echo chamber distribution. This is, however, not really a problem, since we just want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers. \\

@@

When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we can observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors. This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}. Note that for an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for a fraction of completely stubborn actors equal to fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}. This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was already explained for Figure \ref{echo_20_80}. \\

@@

The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher. It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of an initial $20/80$ opinion distribution. The reason for this is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}. When comparing the different network models, these skewed distributions should not be overlooked. The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other. \\

@@

The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, look again different than those for the PR filtering algorithm. The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease. This result is very similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}. The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for a fraction of completely stubborn actors smaller than fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we can start to see a decrease in the formation of echo chambers. The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show a tiny, tiny formation of echo chambers for a fraction of zero stubborn actors (echo chamber size around $1.2$). This very limited, almost non-significant formation of echo chambers for the ER model and the SBM with a low modularity, in combination with the REC filtering algorithm, is also visible in Figure \ref{er_echo_20-80} and Figure \ref{sbm}. For an increasing fraction of the completely stubborn actors, these small echo chambers eventually converge to one, indicating a constant amount of echo chambers over time.\\

@@

So far the formation of echo chambers in case of an unbalanced initial opinion distribution and introducing a fraction of stubborn actors in the system. Lets now discuss the evolution of the opinions in the system. In Figure \ref{ev_op_20_80}, we already saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. The PR filtering algorithm, on the other hand, resulted in a strong decrease of the prevalence of the minority opinion. Are these results affected by introducing stubborn actors? As for the balanced initial conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm. Introducing stubborn actors in the system combined with the PR filtering algorithm, on the other hand, does have an impact in the case of unbalanced initial conditions.\\

@@

Figure \ref{ev_op_20_80_stubb_PR} shows the results for the fraction of opinion A versus time in case of an unbalanced initial opinion distribution and when using the PR filtering algorithm. Data points are presented along with their standard deviation $\sigma$. Again, we compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS) and for each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). \\

@@

If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of an initial unbalanced opinion distribution. Introducing a fraction of stubborn actors changes this scenario. For an increasing fraction of stubborn actors, the decrease of the minority opinion becomes less and less strong. This result is the same for all the different network topologies. If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion. If we however introduce a fraction of stubborn actors in the system, then we will also end up with some stubborn actors that carry the minority opinion. These stubborn actors with the minority opinion will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them. The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{ER}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{WS}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{SBM}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{SBM-WS}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the fraction of nodes with opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with an average degree around $\left<k\right> \sim 10$ and a total number of nodes $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model (ER), $p=0.01$; \textbf{(b)} Watts-Strogatz model (WS) with $K = 10$, $\beta = 0.06$ and $\left<cc\right> \sim 0.55$; \textbf{(c)} Stochastic block model (SBM) with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model (SBM-WS) with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$. Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization.}%more explanation?
\label{ev_op_20_80_stubb_PR}
\end{figure}
\noindent
\textbf{Making all nodes partly stubborn}\\
\newline
Instead of introducing a fraction of completely stubborn actors in the system, as was done before, we will now make all the nodes partly stubborn (so they still have some chance of changing their opinion, they are however less eager to do so). All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
\newline
We will first look at the formation of echo chambers and the evolution of the opinions in the system for an initial $50/50$ opinion distribution, where-after we will investigate the results for an unbalanced initial opinion distribution ($20/80$).

@@

	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_12x10.png}

@@

	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with a total number of nodes $N=10^3$ and an average degree around $\left<k\right> \sim 10$: WS network with $K =10,\ \beta = 0.06$ and $\left<cc\right> \sim 0.55$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $K = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> \sim 0.55$ and $Q \sim 0.9$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q \sim 0.9$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q \sim 0.2$) are depicted; ER network with $p= 0.01$. Each data point is the average of $10^2$ independent simulations.}

@@

Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod} and the data points are shown together with their standard error $SE$. The results look very similar to the case where a fraction of the nodes is completely stubborn and when using the PR filtering algorithm, which is depicted in Figure \ref{echo_vs_frac_complStubb}.\\

@@

The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers, as before, and this is not changed by making the nodes more and more stubborn. The same result is obtained for the stochastic block model with a low modularity (the data points lie again behind those of the ER network).\\

@@

The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$. If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases. This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$. If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease. If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same. In Figure \ref{echo_vs_all_frac_stubb_PR} this is represented by an echo chamber size equal to one for $r=1$.\\

@@

In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution. Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm. Data points are again displayed with their standard error $SE$.\\

@@

We can make similar observations as we did in the ``Making a fraction of the nodes completely stubborn'' paragraph. The results for the REF and REC filtering algorithms look qualitatively similar to each other and differ greatly from the results obtained with the PR filtering algorithm depicted in Figure \ref{echo_vs_all_frac_stubb_PR}. The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases. For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slightly decrease between $r\sim 0.7$ and $r\sim 0.9$. Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers (echo chamber values around 17 for $r\sim 0.9$ to echo chamber values equal to one for $r=1$). The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model, the SBM-WS model on the other hand displays a continuous increase until $r \sim 0.9$. In both case we have again a sharp drop from $r \sim 0.9$ to $r=1$. If $r=1$, the nodes are completely stubborn and thus keep their initial opinion. Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one. It is however remarkable that for very high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is very large and even larger than if the nodes are not stubborn. We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithm in the case of networks with a high clustering coefficient. This is the complete opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes). The reason for this discrepancy is probably of a similar nature as the reason suggested for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Making a fraction of the nodes completely stubborn'' paragraph. If we zoom in on the stochastic block model with a high modularity, we can also see a slight increase in the formation of echo chambers. The increase is however not as strong as for the high clustering networks.\\

@@

The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (again, data points are hidden behind those of the ER network) do not seem to have any dependency on the stubbornness level of the nodes. This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithm that drives this increase in the formation of echo chambers with increasing $r$. % talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??

>>>>

  \caption[Formation of echo chambers versus fraction of completely stubborn actors for the PR and REC filtering algorithms and an initial $20/80$ opinion distribution.]{Formation of echo chambers versus the fraction of completely stubborn actors, fracRes, for an initial $20/80$ opinion distribution. \textbf{(a)} PR filtering algorithm; \textbf{(b)} REC filtering algorithm. The $y$-axis shows the total number of like-minded neighborhoods (sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B) at $t=500$ divided by the total number of like-minded neighborhoods at $t=0$. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$. Each data point is the average of $10^2$ independent simulations. Bars represent the standard error $SE$.}

@@

Figure \ref{echo_vs_frac_complStubb_PR-REC_20-80} shows the formation of echo chambers versus the fraction of completely stubborn actors ($r_i = 1$), fracRes, for an initial $20/80$ opinion distribution (opinion A is the minority opinion, opinion B is the majority opinion). The PR and the REC filtering algorithms are compared. The standard errors $SE$ are small and not visible in the figure. The dotted black line represent the situation in which the number of echo chambers remains constant over time. The interpretation of the $y$-axis is slightly different than before. In previous figures, the echo chamber size was interpreted as the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. Thus we first normalized the echo chambers around opinion A and those around opinion B and than took the average of both. In case of a $20/80$ distribution we have a very skewed echo chamber distribution, as can be seen in Figure \ref{echo_20_80}, with normalized values going to infinity. We still want to have some average of the echo chambers around opinion A and those around opinion B to quantify the total formation of echo chambers in the system, but we want to avoid problems with those infinities. Hence, we opted for a different average: first we calculate the total number of like-minded neighborhoods at $t=500$ (thus the sum of the number of echo chambers around opinion A and the number of echo chambers around opinion B), then we divide that quantity by the total number of like-minded neighborhoods at $t=0$ (hence, we first take the average and than normalize it, instead of normalizing first and then taking the average, as was done before). By taking the average, we neglect the skewed echo chamber distribution. This is, however, not a problem. We want a sense of the total number of echo chambers in the system and how this total number is impacted by the presence of stubborn actors and are less interested in the distribution of those echo chambers. \\

@@

When using the PR filtering algorithm, see Figure \ref{PR_frac_compl_stubb_20-80}, we observe a decrease in the formation of echo chambers with an increasing fraction of stubborn actors. This behavior is similar to the case of an initial $50/50$ opinion distribution in combination with the PR filtering algorithm, as depicted in Figure \ref{echo_vs_frac_complStubb}. For an initial $20/80$ distribution, in combination with the PR filtering algorithm, the formation of echo chambers in the Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity is non-zero (for fracRes $=0$), as can also be observed in Figure \ref{echo_20_80}. This behavior is a consequence of the fact that for an initial $20/80$ opinion distribution, in combination with the PR filtering algorithm, the prevalence of the majority opinion strongly increases, as was explained for Figure \ref{echo_20_80}. \\

@@

The formation of echo chambers for networks with a high clustering and/or a high modularity is slightly higher. It is however not recommended to (quantitatively) compare the formation of echo chambers between the different network models when depicting the results as the average of the echo chambers around opinion A and those around opinion B in case of unbalanced starting conditions. The reason is that this average neglects the skewed echo chamber distributions as depicted in Figure \ref{echo_20_80}. When comparing the different network models, these skewed distributions should not be overlooked. The goal of this figure is to show the dependency of the formation of echo chambers in the network models on the presence of a fraction of stubborn actors and not to (quantitatively) compare the different models against each other. \\

@@

The results for the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb_20-80}, are different than those for the PR filtering algorithm. The high clustering networks (WS and SBM-WS) show an increase in the formation of echo chambers with an increasing fraction of completely stubborn actors, followed by a decrease. This result is similar to the results for an initial $50/50$ distribution in combination with the REC filtering algorithm, depicted in Figure \ref{REC_frac_compl_stubb}. The stochastic block model with a high modularity does not show an increase in the formation of echo chambers but remains flat for fracRes $\leqslant 0.2$; if the fraction of stubborn actors increases further, we see a decrease in the formation of echo chambers. The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity show no significant formation of echo chambers. For an increasing fraction of the completely stubborn actors, the echo chambers eventually converge to one (for all the network models), indicating a constant amount of echo chambers over time.\\

@@

Lets now discuss the evolution of the prevalence of the opinions. In Figure \ref{ev_op_20_80}, we saw that for a system with no stubborn actors the REC filtering algorithm does not break the prevalence of the initial opinions in case of an initial $20/80$ opinion distribution. The PR filtering algorithm, on the other hand, resulted in a decrease of the prevalence of the minority opinion. Are these results affected by introducing stubborn actors? As for the balanced initial conditions, introducing stubborn actors does not break the prevalence of the initial opinions for the REC filtering algorithm. The status quo is maintained over time. Introducing stubborn actors combined with the PR filtering algorithm, on the other hand, does have an impact in case of unbalanced initial conditions.\\

@@

Figure \ref{ev_op_20_80_stubb_PR} shows the results for the prevalence of opinion A versus time in case of unbalanced starting conditions and using the PR filtering algorithm. We compare four different network models (ER, WS, SBM (high and low modularity) and SBM-WS). For each model we compare the results for three different fractions of stubborn actors (fracRes $=0$, fracRes $=0.4$ and fracRes $=0.8$). \\

@@

If there are no stubborn actors, the PR filtering algorithm leads to a significant decrease in the prevalence of the minority opinion in case of unbalanced starting conditions. Introducing a fraction of stubborn actors changes this scenario. For an increasing fraction of stubborn actors, the decrease of the minority opinion becomes less and less strong. This result is valid for all the different network topologies. If there are no stubborn actors, the PR filtering algorithm forces nodes with the minority opinion to switch to the majority opinion. If we introduce a fraction of stubborn actors in the system, then we will end up with stubborn actors that carry the minority opinion. These stubborn actors will not switch to the majority opinion, no matter how hard the PR filtering algorithm pushes them. The consequence is that these stubborn actors with the minority opinion hamper the decrease of that minority opinion.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{ER}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_ER_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{WS}
  	\includegraphics[width=8cm, height=7cm]{fraction_of_opinions_WS_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{SBM}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{SBM-WS}
    \includegraphics[width=8cm, height=7cm]{fraction_of_opinions_SBM-WS_frac_compl_stubb_PR_20-80_8x7.png}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ opinion distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$).]{Evolution of the prevalence of opinion A ($P_\text{A}(t)$) over time for an initial $20/80$ distribution, using the PR filtering algorithm and comparing three different fractions of completely stubborn actors (fracRes $= 0$, fracRes $= 0.4$ and fracRes $= 0.8$). Four different network models are used, each with average degree $\left<k\right> \sim 10$ and $N = 10^3$: \textbf{(a)} Erd\H{o}s-R\'{e}nyi model, $p=0.01$; \textbf{(b)} Watts-Strogatz model with $\left<k\right> = 10$, $\beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; \textbf{(c)} Stochastic block model with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; \textbf{(d)} Stochastic Watts-Strogatz block model with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$. Each curve is the average of $10^2$ independent simulations and data points are shown for every 50 time steps to improve visualization. Bars represent the standard deviation $\sigma$.}%more explanation?
\label{ev_op_20_80_stubb_PR}
\end{figure}
\noindent
\textbf{All actors partly stubborn}\\
\newline
In the second configuration, all the nodes are partly stubborn. They still have some chance of changing their opinion, they are however less eager to do so. All the nodes are given the same stubbornness parameter $r$ and we let this stubbornness parameter vary from $r = 0$ (all the nodes are completely willing to change opinions) to $r=1$ (all the nodes are completely stubborn).\\
\newline
We first study the formation of echo chambers and the evolution of the prevalence of the opinions for balanced starting conditions, where-after we investigate the results for unbalanced starting conditions.

@@

	\includegraphics[width=10cm, height=8cm]{echo_chamber_vs_stubbornness_all_frac_stubb_PR_10x8.png}

@@

	\caption[Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for the PR filtering algorithm and an initial $50/50$ opinion distribution.]{Formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The $y$-axis shows the average of the normalized fraction of nodes with all neighbors having opinion A and the normalized fraction of nodes with all neighbors having opinion B. This value represents the formation of echo chambers in the system (denoted with echo chamber size in the figure). Results are shown for four different network models, each with $N=10^3$ and average degree $\left<k\right> \sim 10$: WS network with $\left<k\right> =10,\ \beta = 0.06$ and $\left<cc\right> = 0.557 \pm 0.007$; SBM-WS network with 10 communities of size 100, each community has the following WS parameters: $\left<k\right> = 10,\ \beta = 0.01$ and $p_{\text{add}} = 0.001$, $\left<cc\right> = 0.553 \pm 0.004$ and $Q = 0.909 \pm 0.004$; SBM network with 10 communities of size 100, a high modularity case ($p_{\text{cl}} = 0.1,\ p_{\text{add}} = 0.001,\ Q = 0.908 \pm 0.005$) and a low modularity case ($p_{\text{cl}} = 0.03,\ p_{\text{add}} = 0.008,\ Q = 0.220 \pm 0.007$) are depicted; ER network with $p= 0.01$. Each data point is the average of $10^2$ independent simulations. Bars represent the standard error $SE$.}

@@

Figure \ref{echo_vs_all_frac_stubb_PR} shows the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, for an initial $50/50$ opinion distribution and using the PR filtering algorithm. The interpretation of the $y$-axis is again as explained in Section \ref{echoVSmod}. The results are similar to the case where a fraction of the nodes is completely stubborn combined with the PR filtering algorithm (Fig. \ref{echo_vs_frac_complStubb}).\\

@@

The Erd\H{o}s-R\'{e}nyi model has no appearance of echo chambers. This is not changed by making the nodes more and more stubborn. The same result is obtained for the stochastic block model with a low modularity (the data points lie behind those of the ER network).\\

@@

The stochastic block model with a high modularity has a (small) formation of echo chambers for $r=0$. If the nodes become more and more convinced of their own opinion ($r$ increases), the formation of echo chambers decreases. This is especially visible for the high clustering networks (WS and SBM-WS), which have the highest formation of echo chambers for $r=0$. If $r$ increases and the nodes become more and more convinced of their own beliefs, which makes them less and less willing to change that belief, the formation of echo chambers starts to decrease. If $r=1$, the nodes retain their initial opinion and the number of echo chambers before and after the time evolution remains the same. This situation is represented by the black dotted line.\\

@@

In Figure \ref{echo_vs_all_frac_stubb_REF-REC} the formation of echo chambers versus the stubbornness parameter of all the nodes, $r$, is depicted for an initial $50/50$ opinion distribution. Figure \ref{REF_all_frac_stubb} shows the results for the REF filtering algorithm and Figure \ref{REC_all_frac_stubb} shows those for the REC filtering algorithm.\\

@@

We can make similar observations as we did in the ``Completely stubborn actors'' paragraph. The results for the REF and REC filtering algorithms are qualitatively similar to each other and differ from the results obtained with the PR filtering algorithm (Fig. \ref{echo_vs_all_frac_stubb_PR}). The formation of echo chambers in the high clustering networks (WS and SBM-WS) increases if the stubbornness of the nodes $r$ increases. For the REF filtering algorithm, this increases reaches a maximum around $r \sim 0.7$ and then starts to slowly decrease between $r\sim 0.7$ and $r\sim 0.9$. Between $r \sim 0.9$ and $r=1$ there is a steep drop in the formation of echo chambers. The REC filtering algorithm also seems to have a maximum around $r \sim 0.7$ for the WS model. The SBM-WS model, on the other hand, displays a continuous increase until $r \sim 0.9$. In both case we have a sharp drop from $r \sim 0.9$ to $r=1$. If $r=1$, the nodes are completely stubborn and thus keep their initial opinion. Hence, the number of like minded neighborhoods does not change over time, which gives a normalized number of echo chambers equal to one. This is represented by the black dotted line. It is remarkable that for high levels of the stubbornness of the nodes $r$, up to $r \sim 0.9$, the formation of echo chambers is large and even larger than if the nodes are not stubborn. We can conclude that making the nodes partly stubborn enhances the formation of echo chambers when using the REC/REF filtering algorithms in case of networks with a high clustering coefficient. This is the opposite of the case where the PR filtering algorithm is used (which shows a decrease in the formation of echo chambers with an increasing stubbornness of the nodes). The reason for this discrepancy is probably of a similar nature as for this discrepancy in case of introducing a fraction of completely stubborn actors, see the ``Completely stubborn actors'' paragraph. If we zoom in on the stochastic block model with a high modularity, we see a slight increase in the formation of echo chambers. The increase is however not as strong as for the high clustering networks.\\

@@

The Erd\H{o}s-R\'{e}nyi network and the stochastic block model with a low modularity (data points are hidden behind those of the ER network) do not have any dependency on the stubbornness level of the nodes. This seems to suggest that it is mainly the interplay between the high clustering, and to a lesser extent a clear community structure (high modularity), and the REF/REC filtering algorithms that drives this increase in the formation of echo chambers with increasing $r$. % talk about fact that you go from initial random 50/50 distribution to opinion clustered 50/50 distribution??
